---
title: "Module 6 | R" 
author: "Ryan S Dunn"
date: "11/30/2021"
output: pdf_document
---
## Data Science Using Python and R: Chapter 10 - Page 149: Questions #11, 12, 13, & 14

### 11. Input and standardize both the training and test data sets.

```{r}
#import the data into R
wine_train <- read.csv("/Users/ryan_s_dunn/Documents/USD MS-ADS/Applied Data Mining 502/Module 6/datasets/white_wine_training")
wine_test <- read.csv("/Users/ryan_s_dunn/Documents/USD MS-ADS/Applied Data Mining 502/Module 6/datasets/white_wine_test")
```

```{r}
#subset and standardize the training and test data sets
train_subset <- subset(wine_train, select = c("alcohol","sugar"))
test_subset <- subset(wine_test, select = c("alcohol","sugar"))

#standardize training 
train_z <- as.data.frame(scale(train_subset))
colnames(train_z) <- c("alcohol_z","sugar_z")

#standardize test
test_z <- as.data.frame(scale(test_subset))
colnames(test_z) <- c("alcohol_z", "sugar_z")

```

### 12. Run k‐means clustering on the training data set, using two clusters.

```{r}
#run the k-means clustering algorithm on training data 
kmeans01 <- kmeans(train_z, centers = 2)
```

```{r}
#save the cluster memebership of each records as its own variable
cluster <- as.factor(kmeans01$cluster)
```

```{r}
#seperate records into two groups for membership
cluster1 <- train_z[ which(cluster == 1), ]
cluster2 <- train_z[ which(cluster == 2), ]
```
### 13. Give the mean of each variable within each cluster and use the means to identify a “Dry wines” and a “Sweet wines” cluster.

#### Summary statistics for Cluster 1 (training)
```{r}
summary(cluster1)
```

#### Summary statistics for Cluster 2 (training)
```{r}
summary(cluster2)
```
### 14. Validate the clustering results by running k‐means clustering on the test data set, using two clusters, and identifying a “Dry wines” and a “Sweet wines” cluster.

```{r}
#run the k-means clustering algorithm on test data 
kmeans01_test <- kmeans(train_z, centers =2)
```

```{r}
#save the cluster memebership of each records as its own variable
cluster_test <- as.factor(kmeans01_test$cluster)
```

```{r}
#seperate records into two groups for membership
cluster1_test <- test_z[ which(cluster == 1), ]
cluster2_test <- test_z[ which(cluster == 2), ]
```

#### Summary statistics for Cluster 1 (test)

```{r}
summary(cluster1_test)
```

#### Summary statistics for Cluster 2 (test)

```{r}
summary(cluster2_test)
```






